---
author: "Mark Torrance"
date: "17 May 2018"
output: html_document
---

#Specifying Contrasts for linear model
At the moment my motivation for making this available is so that someone can tell me if I'm doing what I want to do.

Get ```tidyverse``` for data wrangling and ```lme4``` for modelling and load some data that I prepared earlier.

```{r echo = T, message = F}
library(tidyverse)
library(lme4)

load("rti_narrative_data.Rda")
```


This loads a data frame called ```Df``` that looks like this:

```{r, eval = T, echo = F}
summary(Df)
```

```T1234``` is a factor representing tests at four different time points.

```Condition``` comprises two groups, ```comp``` and ```int```

I am constructing comparisons as follows.

Create a "flat" single factor representing all cells in the design.

```{r, eval = T}
Df <- Df %>% 
  mutate(cond_by_task = interaction(T1234,Condition))
summary(Df$cond_by_task)
```

Then define some contrasts:

Main effect of group

```{r, eval = T}
cond <- cbind(c(-1,-1,-1,-1,1,1,1,1))
colnames(cond) <- c(' group')
```


Main effects of time with separate contrasts giving slope, averaged across groups, between T1 and T2, T2 and T3, and T3 and T4.

```{r, eval = T}
time <- cbind(rep(c(-1,1,0,0),2),
             rep(c(0,-1,1,0),2),
             rep(c(0,0,-1,1),2))
colnames(time) <- c(' T12',' T23',' T34')
```

Interaction effects. This asks whether there the slopes for the two groups differ, looking separately at slopes between T1 and T2, T2 and T3, and T3 and T4.

```{r, eval = T}
inter <- cbind(c(1,-1,0,0,-1,1,0,0),
              c(0,1,-1,0,0,-1,1,0),
              c(0,0,1,-1,0,0,-1,1))
colnames(inter) <- c(' T12:group',' T23:group',' T34:group')
```


Then put these all together and assign them to the ```cond_by_task``` factor.

```{r, eval = T}
conts <- cbind(cond,time,inter)
contrasts(Df$cond_by_task, how.many = ncol(conts)) <- conts
contrasts(Df$cond_by_task)
```


You don't need an intercept because lmer gives you it for free (and won't allow you to use your own contrast but specify no intercept in the model).

Then run the model and get the coefficients.

```{r, eval = T}
m <- lmer(score ~ cond_by_task + (1|Subno), 
          data = Df)

summary(m)$coef
```

***

So, the question is: Is this set of contrasts giving me what I've claimed it's giving me? If not, what am I doing wrong?
